\documentclass[ebook,11pt,article]{memoir}
\usepackage{geometry}  % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}  % ... or a4paper or a5paper or ... 
%\geometry{landscape}  % Activate for for rotated page geometry

\pagestyle{myheadings}
%%% from std.tex
%\usepackage[american]
%           {babel}        % needed for iso dates
%\usepackage[iso,american]
%           {isodate}      % use iso format for dates
\usepackage[final]
           {listings}     % code listings
%\usepackage{longtable}    % auto-breaking tables
%\usepackage{ltcaption}    % fix captions for long tables
\usepackage{relsize}      % provide relative font size changes
%\usepackage{textcomp}     % provide \text{l,r}angle
\usepackage{underscore}   % remove special status of '_' in ordinary text
%\usepackage{parskip}      % handle non-indented paragraphs "properly"
%\usepackage{array}        % new column definitions for tables
\usepackage[normalem]{ulem}
\usepackage{enumitem}
\usepackage{color}        % define colors for strikeouts and underlines
%\usepackage{amsmath}      % additional math symbols
%\usepackage{mathrsfs}     % mathscr font
\usepackage[final]{microtype}
%\usepackage{multicol}
\usepackage{xspace}
%\usepackage{lmodern}
\usepackage[T1]{fontenc} % makes tilde work! and is better for umlauts etc.
%\usepackage[pdftex, final]{graphicx}
\usepackage[pdftex,
%            pdftitle={C++ International Standard},
%            pdfsubject={C++ International Standard},
%            pdfcreator={Richard Smith},
            bookmarks=true,
            bookmarksnumbered=true,
            pdfpagelabels=true,
            pdfpagemode=UseOutlines,
            pdfstartview=FitH,
            linktocpage=true,
            colorlinks=true,
            linkcolor=blue,
            plainpages=false
           ]{hyperref}
%\usepackage{memhfixc}     % fix interactions between hyperref and memoir
%\usepackage[active,header=false,handles=false,copydocumentclass=false,generate=std-gram.ext,extract-cmdline={gramSec},extract-env={bnftab,simplebnf,bnf,bnfkeywordtab}]{extract} % Grammar extraction
%
\renewcommand\RSsmallest{5.5pt}  % smallest font size for relsize


%%%% reuse all three from std.tex:
\input{macros}
\input{layout}
\input{styles}

%%%%

\newcommand{\papernumber}{P0052R4}
\newcommand{\paperdate}{2017-06-15}

\markboth{\papernumber{} \paperdate{}}{\papernumber{} \paperdate{}}

\title{\papernumber{} - Generic Scope Guard and RAII Wrapper for the Standard Library}
\author{Peter Sommerlad and Andrew L. Sandoval\\with contributions by Eric Niebler and Daniel Kr\"ugler}
\date{\paperdate}                        % Activate to display a given date or no date
\setsecnumdepth{subsection}

\begin{document}
\maketitle
\begin{tabular}[t]{|l|l|}\hline 
Document Number: \papernumber &   (update of N4189, N3949, N3830, N3677)\\\hline
Date: & \paperdate \\\hline
Project: & Programming Language C++\\\hline 
Audience: & LWG/LEWG\\\hline
\end{tabular}

\chapter{History}
\section{Changes from P0052R3}
\begin{itemize}
\item Take new section numbering of the standard working paper into account.
\item require noexcept of \tcode{f()} for \tcode{scope_exit} and \tcode{scope_fail} explicitly
\item implementation could be tested with C++17 compiler and class template constructor argument deduction thus the paper no longer claims help or not being sure.
\end{itemize}


\section{Changes from P0052R2}
\begin{itemize}
\item Take into account class template ctor argument deduction. However, I recommend keeping the factories for LFTS 3 to allow for C++14 implementations. At the time of this writing, I do not have a working C++17 compliant compiler handy to run corresponding test cases without the factories. However, there is one factory function make_unique_checked that needs to stay, because it addresses a specific but seemingly common use-case.
\item Since \tcode{scope_success} is a standard library class that has a possible throwing destructor section [res.on.exception.handling] must be adjusted accordingly.
\item The lack of factories for the classes might require explicit deduction guides, but I need help to specify those accordingly since I do not have a working C++17 compiler right at hand to test it.
\end{itemize}


\section{Changes from P0052R1}
The Jacksonville LEWG, especially Eric Niebler gave splendid input in how to improve the classes in this paper. I (Peter) follow Eric's design in specifying scope_exit as well as unique_resource in a more general way.
\begin{itemize}
\item Provide \tcode{scope_fail} and \tcode{scope_success} as classes. However, we may even hide all of the scope guard types and just provide the factories.
\item safe guard all classes against construction errors, i.e., failing to copy the deleter/exit-function, by calling the passed argument in the case of an exception, except for scope_success.
\item relax the requirements for the template arguments.
\end{itemize}
Special thanks go to Eric Niebler for providing several incarnations of an implementation that removed previous restrictions on template arguments in an exception-safe way (Eric: \textit{"This is HARD."}). To cite Eric again: \textit{"Great care must be taken when move-constructing or
move-assigning unique_resource objects to ensure that there is always
exactly one object that owns the resource and is in a valid,
Destructible state."}
Also thanks to Axel Naumann for presenting in Jacksonville and to Axel, Eric, and Daniel Kr\"ugler for their terrific work on wording improvements.

\section{Changes from P0052R0}
In Kona LWG gave a lot of feedback and especially expressed the desire to simplify the constructors and specification by only allowing \emph{nothrow-copyable} \tcode{RESOURCE} and \tcode{DELETER} types. If a reference is required, because they aren't, users are encouraged to pass a \tcode{std::ref/std::cref} wrapper to the factory function instead.
\begin{itemize}
\item Simplified constructor specifications by restricting on nothrow copyable types. Facility is intended for simple types anyway. It also avoids the problem of using a type-erased \tcode{std::function} object as the deleter, because it could throw on copy.
\item Add some motivation again, to ease review and provide reason for specific API issues.
\item Make "Alexandrescu's" "declarative" scope exit variation employing \tcode{uncaught_exceptions()} counter optional factories to chose or not.
\item propose to make it available for standalone implementations and add the header \tcode{<scope>} to corresponding tables.
\item editorial adjustments
\item re-established \tcode{operator*} for \tcode{unique_resource}.
\item overload of \tcode{make_unique_resource} to handle \tcode{reference_wrapper} for resources. No overload for reference-wrapped deleter functions is required, because \tcode{reference_wrapper} provides the call forwarding.
\end{itemize}
\section{Changes from N4189}
\begin{itemize}
\item Attempt to address LWG specification issues from Cologne (only learned about those in the week before the deadline from Ville, so not all might be covered).
\begin{itemize}
\item specify that the exit function must be either no-throw copy-constructible, or no-throw move-constructible, or held by reference. Stole the wording and implementation from unique_ptr's deleter ctors.
\item put both classes in single header \tcode{<scope>}
\item specify factory functions for Alexandrescu's 3 scope exit cases for \tcode{scope_exit}. Deliberately did't provide similar things for \tcode{unique_resource}.
\end{itemize}
\item remove lengthy motivation and example code, to make paper easier digestible.
\item Corrections based on committee feedback in Urbana and Cologne.
\end{itemize}

%TODO AM: I want to see a proper synopsis with summary, and separate class definition.
%TODO DK: Have we discussed why make_scoped does remove_reference but not cv?
%TODO: single header or merge with <utility>
%DK: re make_scope_exit function: Wondering why there's remove_reference instead of decay. Means you can't construct scope_exit from a function, and means function object member may be cv-qualified, which can affect overload resolution. Is this designed?
%TODO AM: We rarely put the function definition in the class definition. Rampant use of unqualified noexcept when wrapping unknown type seems totally bogus. For make_scoped_exit, deduced return type doesn't play nicely with SFINAE and other such things. We discussed this in Urbana.
%TODO DK: Agree on the last point, but for unconditional noexcept, there's a documented requirement that move construction shall not throw an exception. So this seems to be by design.
%TODO noexcept - moving - non-throwing
% TODO Alexandrescu: UncaughtExceptionCount... ScopeGuardForNewExeption(slide photo!)
% github.com/facebook/folly
% github.com/panaseleus/stack_unwinding

\section{Changes from N3949}
\begin{itemize}
\item renamed \tcode{scope_guard} to \tcode{scope_exit} and the factory to \tcode{make_scope_exit}. Reason for make_ is to teach users to save the result in a local variable instead of just have a temporary that gets destroyed immediately. Similarly for unique resources, \tcode{unique_resource}, \tcode{make_unique_resource} and \tcode{make_unique_resource_checked}.
\item renamed editorially \tcode{scope_exit::deleter} to \tcode{scope_exit::exit_function}.
\item changed the factories to use forwarding for the \tcode{deleter}/\tcode{exit_function} but not deduce a reference.
\item get rid of \tcode{invoke}'s parameter and rename it to \tcode{reset()} and provide a \tcode{noexcept} specification for it.
\end{itemize}


\section{Changes from N3830}
\begin{itemize}
\item rename to \tcode{unique_resource_t} and factory to \tcode{unique_resource}, resp. \tcode{unique_resource_checked}
\item provide scope guard functionality through type \tcode{scope_guard_t} and \tcode{scope_guard} factory
\item remove multiple-argument case in favor of simpler interface, lambda can deal with complicated release APIs requiring multiple arguments.
\item make function/functor position the last argument of the factories for lambda-friendliness.

\end{itemize}

\section{Changes from N3677}
\begin{itemize}
\item Replace all 4 proposed classes with a single class covering all use cases, using variadic templates, as determined in the Fall 2013 LEWG meeting.
\item The conscious decision was made to name the factory functions without "make", because they actually do not allocate any resources, like \tcode{std::make_unique} or \tcode{std::make_shared} do
\end{itemize}

\chapter{Introduction}
The Standard Template Library provides RAII (resource acquisition is initialization) classes for managing pointer types, such as \tcode{std::unique_ptr} and \tcode{std::shared_ptr}.  This proposal seeks to add a two generic RAII wrappers classes which tie zero or one resource to a clean-up/completion routine which is bound by scope, ensuring execution at scope exit (as the object is destroyed) unless released early or in the case of a single resource: executed early or returned by moving its value.

\chapter{Acknowledgements}
\begin{itemize}
\item This proposal incorporates what Andrej Alexandrescu described as scope_guard long ago and explained again at C++ Now 2012 (%\url{
%https://onedrive.live.com/view.aspx?resid=F1B8FF18A2AEC5C5!1158&app=WordPdf&wdo=2&authkey=!APo6bfP5sJ8EmH4}
).
\item This proposal would not have been possible without the impressive work of Peter Sommerlad who produced the sample implementation during the Fall 2013 committee meetings in Chicago.  Peter took what Andrew Sandoval produced for N3677 and demonstrated the possibility of using C++14 features to make a single, general purpose RAII wrapper capable of fulfilling all of the needs presented by the original 4 classes (from N3677) with none of the compromises.
\item Gratitude is also owed to members of the LEWG participating in the Fall 2015(Kona),Fall 2014(Urbana), February 2014 (Issaquah) and Fall 2013 (Chicago) meeting for their support, encouragement, and suggestions that have led to this proposal.
\item Special thanks and recognition goes to OpenSpan, Inc. (http://www.openspan.com) for supporting the production of this proposal, and for sponsoring Andrew L. Sandoval's first proposal (N3677) and the trip to Chicago for the Fall 2013 LEWG meeting. \emph{Note: this version abandons the over-generic version from N3830 and comes back to two classes with one or no resource to be managed.}
\item Thanks also to members of the mailing lists who gave feedback. Especially Zhihao Yuan, and Ville Voutilainen.
\item Special thanks to Daniel Kr\"ugler for his deliberate review of the draft version of this paper (D3949).
\end{itemize}
\newpage
\chapter{Motivation}
While \tcode{std::unique_ptr} can be (mis-)used to keep track of general handle types with a user-specified deleter it can become tedious and error prone. Further argumentation can be found in previous papers. Here are two examples using  \tcode{<cstdio>}'s \tcode{FILE *} and POSIX\tcode{<fcntl.h>}'s and \tcode{<unistd.h>}'s \tcode{int} file handles. 

\begin{codeblock}
void demonstrate_unique_resource_with_stdio() {
  const std::string filename = "hello.txt";
  { auto file=make_unique_resource(::fopen(filename.c_str(),"w"),&::fclose);
    ::fputs("Hello World!\n", file.get());
    ASSERT(file.get()!= NULL);
  }
  { std::ifstream input { filename };
    std::string line { };
    getline(input, line);
    ASSERT_EQUAL("Hello World!", line);
    getline(input, line);
    ASSERT(input.eof());
  }
  ::unlink(filename.c_str());
  {
    auto file = make_unique_resource_checked(::fopen("nonexistingfile.txt", "r"), 
                (FILE*) NULL, &::fclose);
    ASSERT_EQUAL((FILE*)NULL, file.get());
  }
}
void demontrate_unique_resource_with_POSIX_IO() {
  const std::string filename = "./hello1.txt";
  { auto file=make_unique_resource(::open(filename.c_str(),
                     O_CREAT|O_RDWR|O_TRUNC,0666), &::close);
    ::write(file.get(), "Hello World!\n", 12u);
    ASSERT(file.get() != -1);
  }
  { std::ifstream input { filename };
    std::string line { };
    getline(input, line);
    ASSERT_EQUAL("Hello World!", line);
    getline(input, line);
    ASSERT(input.eof());
  }
  ::unlink(filename.c_str());
  {
    auto file = make_unique_resource_checked(::open("nonexistingfile.txt", 
                       O_RDONLY), -1, &::close);
    ASSERT_EQUAL(-1, file.get());
  }
}\end{codeblock}

We refer to Andrej Alexandrescu's well-known many presentations as a motivation for \tcode{scope_exit}, \tcode{scope_fail}, and \tcode{scope_success}. Here is a brief example on how to use the 3 proposed factories. 
\begin{codeblock}
void demo_scope_exit_fail_success(){
  std::ostringstream out{};
  auto lam=[&]{out << "called ";};
  try{
    auto v=make_scope_exit([&]{out << "always ";});
    auto w=make_scope_success([&]{out << "not ";}); // not called
    auto x=make_scope_fail(lam); // called
    throw 42;
  }catch(...){
    auto y=make_scope_fail([&]{out << "not ";}); // not called
    auto z=make_scope_success([&]{out << "handled";}); // called
  }
  ASSERT_EQUAL("called always handled",out.str());
}
\end{codeblock}


\chapter{Impact on the Standard}
This proposal is a pure library extension. A new header, \tcode{<scope>} is proposed, but it does not require changes to any standard classes or functions. Since it proposes a new header, no feature test macro seems required. It does not require any changes in the core language, and it has been implemented in standard C++ conforming to \removed{C++14, resp. draft} C++17. Depending on the timing of the acceptance of this proposal, it might go into a library fundamentals TS under the namespace std::experimental or directly in the working paper of the standard.

\chapter{Design Decisions}
\section{General Principles}
The following general principles are formulated for \tcode{unique_resource}, and are valid for \tcode{scope_exit} correspondingly.
\begin{itemize}
\item Transparency - It should be obvious from a glance what each instance of a \tcode{unique_resource} object does.  By binding the resource to it's clean-up routine, the declaration of \tcode{unique_resource} makes its intention clear.
\item Resource Conservation and Lifetime Management - Using \tcode{unique_resource} makes it possible to "allocate it and forget about it" in the sense that deallocation is always accounted for after the \tcode{unique_resource} has been initialized.
\item Exception Safety - Exception unwinding is one of the primary reasons that \tcode{unique_resource} and \tcode{scope_exit}/\tcode{scope_fail} are needed. Therefore, the specification asks for strong safety guarantee when creating and moving the defined types, making sure to call the deleter/exit function if such attempts fail.
\item Flexibility - \tcode{unique_resource} is designed to be flexible, allowing the use of lambdas or existing functions for clean-up of resources. 
\end{itemize}

\section{Prior Implementations}
Please see N3677 from the May 2013 mailing (or http://www.andrewlsandoval.com/scope_exit/) for the previously proposed solution and implementation.  Discussion of N3677 in the (Chicago) Fall 2013 LEWG meeting led to the creation of \tcode{unique_resource} and \tcode{scope_exit} with the general agreement that such an implementation would be vastly superior to N3677 and would find favor with the LEWG.  Professor Sommerlad produced the implementation backing this proposal during the days following that discussion.

N3677 has a more complete list of other prior implementations.

N3830 provided an alternative approach to allow an arbitrary number of resources which was abandoned due to LEWG feedback 

The following issues have been discussed by LEWG already:
\begin{itemize}
\item \textit{Should there be a companion class for sharing the resource \tcode{shared_resource} ?  (Peter thinks no. Ville thinks it could be provided later anyway.) } LEWG: NO.
\item \textit{Should \tcode{~scope_exit()} and \tcode{unique_resource::invoke()} guard against deleter functions that throw with \tcode{try{ deleter(); }catch(...){}} (as now) or not?} LEWG: NO, but provide noexcept in detail.
\item \textit{Does \tcode{scope_exit} need to be move-assignable? } LEWG: NO.
\item Should we make the regular constructor of the scope guard templates private and friend the factory function only? This could prohibit the use as class members, which might sneakily be used to create "destructor" functionality by not writing a destructor by adding a \tcode{scope_exit} member variable.
\\
\emph{It seems C++17's class template constructor argument deduction makes the need for most of the factory functions obsolete and thus this question is no longer relevant. However, I recommend keeping the factories for the LFTS-3 if accepted to allow backporting to C++14.}
\item Should the scope guard classes be move-assignable? Doing so, would enable/ease using them as class members. I do not think this use is good, but may be someone can come up with a use case for that. 
\\
\emph{LEWG already answered that once with NO, but you never know if people change their mind again.}
\end{itemize}

The following issues have been recommended by LWG already:
\begin{itemize}
\item Make it a facility available for free-standing implementations in a new header \tcode{<scope>} (\tcode{<utility>} doesn't work, because it is not available for free-standing implementations)
\end{itemize}

\newpage
\section{Open Issues (to be) Discussed by LEWG / LWG}
Since a lot of the design changed due to Eric Niebler's implementation taking great effort to provide strong exception guarantee, another round through LEWG and LWG seems to be required.
\begin{itemize}
\item which "callable" definition in the standard should be applied (call expression (as it is now) or via INVOKE (\tcode{is_callable_v<EF\&>}). 
\emph{IMHO call expression is fine, since everything is about side-effects and we never return a useful value from any of the function objects.}
\item Should we provide a non-explicit conversion operator to \tcode{R} in \tcode{unique_resource<R,D>} ? Last time people seem to have been strongly against, however, it would make the use of unique_resource much easier in contexts envisioned by author Andrew Sandoval. Please re-visit, since it is omitted here.
\end{itemize}

\newpage
\chapter{Technical Specifications}
The following formulation is based on inclusion to the draft of the C++ standard. However, if it is decided to go into the Library Fundamentals TS, the position of the texts and the namespaces will have to be adapted accordingly, i.e., instead of namespace \tcode{std::} we suppose namespace \tcode{std::experimental::}. 

A very recent draft of the standard already has the requested change below that was suggested by Daniel Kr\"ugler:

\section{Adjust 20.5.4.8 Other functions [res.on.functions]}
Since \tcode{~scope_success()} might throw an exception and we can not specify that in a required behavior clause, we need to allow doing so for the standard library's normative remarks section as well.

In section 20.5.4.8 Other functions [res.on.functions] modify p2 item (2.4) as follows by adding "or \remarks"

\begin{description}
\item[\textit{(2.4)} ---]
if any replacement function or handler function or destructor operation exits via an exception,
unless specifically allowed
in the applicable
\required \added{or \remarks}
paragraph.
\end{description}

However the following adjustment is missing, since the standard library promises that all library classes won't throw on destruction:

\section{Adjust 20.5.5.12 Restrictions on exception handling [res.on.exception.handling]}

Change paragraph 3 as follows:

\pnum
Destructor operations defined in the \Cpp standard library
shall not throw exceptions\added{ unless explicitly specified that they could}.
Every destructor\added{ without an exception specification} in the \Cpp standard library shall behave as if it had a
non-throwing exception specification.




\section{Header}
In section 20.5.1.1 Library contents [contents] add an entry to table 16 (cpp.library.headers) for the new header \tcode{<scope>}. Because of the new header, there is no need for a feature test macro.

In section 20.5.1.3 Freestanding implementations [compliance] add an extra row to table 19 (cpp.headers.freestanding) and 
in section [utilities.general] add the same extra row to table 34 (util.lib.summary)
%%TODO clearer specification 
\begin{table}[htb]
\caption{table 19 and table 34}
\begin{center}
\begin{tabular}{|lcl|}
\hline
&Subclause & Header\\
\hline
23.nn &Scope Guard Support & \tcode{<scope>}\\
\hline
\end{tabular}
\end{center}
\label{utilities}
\end{table}%

\section{Additional sections}
Add a a new section to chapter 23 introducing the contents of the header \tcode{<scope>}.

%\rSec1[utilities.scope]{Scope Guard}
\section{Scope guard support [scope]}
This subclause contains infrastructure for a generic scope guard and RAII (resource acquisition is initialization) resource wrapper.\\
\\
\subsection{Header \tcode{<scope>} synopsis}


\begin{codeblock}
namespace std {
namespace experimental {
template <class EF>
class scope_exit;
template <class EF>
class scope_fail;
template <class EF>
class scope_success;


template<class R,class D>
class unique_resource;

// special factory function
template<class R,class D, class S=R>
unique_resource<decay_t<R>, decay_t<D>>
make_unique_resource_checked(R&& r, S const& invalid, D&& d)
noexcept(is_nothrow_constructible_v<decay_t<R>, R> &&
         is_nothrow_constructible_v<decay_t<D>, D>);

// optional factory functions (obsoleted by C++17 class template ctor argument deduction)
template <class EF>
scope_exit<decay_t<EF>> make_scope_exit(EF&& exit_function) ;
template <class EF>
scope_fail<decay_t<EF>> make_scope_fail(EF&& exit_function) ;
template <class EF>
scope_success<decay_t<EF>> make_scope_success(EF&& exit_function) ;

template<class R,class D>
unique_resource<decay_t<R>, decay_t<D>>
make_unique_resource( R&&  r, D&& d) 
noexcept(is_nothrow_constructible_v<decay_t<R>, R> &&
         is_nothrow_constructible_v<decay_t<D>, D>);

template<class R,class D>
unique_resource<R&, decay_t<D>>
make_unique_resource( reference_wrapper<R>  r, D&& d) 
noexcept(is_nothrow_constructible_v<decay_t<D>, D>);

}}
\end{codeblock}

\pnum
The header  \tcode{<scope>} defines the class templates \tcode{scope_exit}, \tcode{scope_fail}, \tcode{scope_success}, \tcode{unique_resource} and
the factory function template \tcode{make_unique_resource_checked()}.
\begin{note}
For backporting possibilities I recommend that the following factory functions are kept, even they are made obsolete by C++17's constructor template argument deduction:
\tcode{make_scope_exit()}, \tcode{make_scope_success()}, \tcode{make_scope_fail()}, and
\tcode{make_unique_resource()}.
\end{note}

\pnum 
The class templates \tcode{scope_exit}, \tcode{scope_fail}, and \tcode{scope_success} define\emph{ scope guards} that wrap a function object to be called on their destruction.

\pnum
The following clauses describe the class templates \tcode{scope_exit}, \tcode{scope_fail}, and \tcode{scope_success}. In each clause, the name \tcode{\textit{scope_guard}} denotes either of these class templates. In description of class members \tcode{\textit{scope_guard}} refers to the enclosing class.

\newpage
%\rSec2[scope.scope_guard]{Scope guard class templates}}
\subsection {Scope guard class templates [scope.scope_guard]}

\begin{codeblock}
template <class EF>
class @\tcode{\textit{scope_guard}}@ {
public:
  template<class EFP>
  explicit @\tcode{\textit{scope_guard}}@(EFP&& f) ;
  @\tcode{\textit{scope_guard}}@(@\tcode{\textit{scope_guard}}@&& rhs) ;
  ~@\tcode{\textit{scope_guard}}@() noexcept(@\seebelow@);
  void release() noexcept;

  @\tcode{\textit{scope_guard}}@(const @\tcode{\textit{scope_guard}}@&)=delete;
  @\tcode{\textit{scope_guard}}@& operator=(const @\tcode{\textit{scope_guard}}@&)=delete;
  @\tcode{\textit{scope_guard}}@& operator=(@\tcode{\textit{scope_guard}}@&&)=delete;
private:
  EF exit_function;    // exposition only
  bool execute_on_destruction{true}; //exposition only
  int  uncaught_on_creation{uncaught_exceptions()}; // exposition only
};

\end{codeblock}
\pnum
\begin{note}
\tcode{scope_exit} is meant to be a general-purpose scope guard that calls its exit function when a scope is exited. The class templates \tcode{scope_fail} and \tcode{scope_success} share the \tcode{scope_exit}'s interface, only the situation when the exit function is called differs. These latter two class templates memorize the value of \tcode{uncaught_exceptions()} on construction and in the case of \tcode{scope_fail} call the exit function on destruction, when \tcode{uncaught_exceptions()} at that time returns a greater value, in the case of \tcode{scope_success} when \tcode{uncaught_exceptions()} on destruction returns the same or a lesser value.\\
\begin{example}
\begin{codeblock}
void grow(vector<int>&v){
	scope_success guard([]{ cout << "Good!" << endl; });
	v.resize(1024);
}
\end{codeblock}
\end{example}
\end{note}


\pnum
\requires
Template argument \tcode{EF} shall be a function object type ([function.objects]), 
lvalue reference to function, or 
lvalue reference to function object type.
If EF is an object type, it  shall satisfy
the requirements of \tcode{Destructible} (Table~27
%\ref{destructible}
). 
%\tcode{is_callable<EF\&>::value} is \tcode{true}.
Given an lvalue \tcode{f} of type \tcode{EF}, the expression \tcode{f()} shall be well formed and shall have well-defined behavior.
The constructor arguments \tcode{f} in the following constructors shall be a function object (23.14)[function.objects], 
lvalue reference to function, or 
lvalue reference to function object.


\begin{itemdecl}
template<class EFP>
explicit
scope_exit(EFP&& f) ;
\end{itemdecl}

\begin{itemdescr}
%\pnum
\remarks This constructor shall not participate in overload resolution unless \tcode{is_constructible_v<EF,EFP>} is \tcode{true}.

\pnum
\requires Given an lvalue \tcode{f} of type \tcode{EFP}, the expression \tcode{f()} shall be well formed and shall have well-defined behavior and not throw an exception.


\pnum
\effects If \tcode{EFP} is not a lvalue-reference type and \tcode{is_nothrow_constructible_v<EF,EFP>} is \tcode{true}, initialize  \tcode{exit_function} with \tcode{move(f)} otherwise initialize \tcode{exit_function} with \tcode{f}. If the copying throws an exception, calls \tcode{f()}.

\pnum
\throws Any exception thrown during the copying of \tcode{f}.

\end{itemdescr}


\begin{itemdecl}
template<class EFP>
explicit
scope_fail(EFP&& f) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\remarks This constructor shall not participate in overload resolution unless \tcode{is_constructible_v<EF,EFP>} is \tcode{true}.

\pnum
\requires Given an lvalue \tcode{f} of type \tcode{EFP}, the expression \tcode{f()} shall be well formed and shall have well-defined behavior and not throw an exception.

\pnum
\effects If \tcode{EFP} is not a lvalue-reference type and \tcode{is_nothrow_constructible_v<EF,EFP>} is \tcode{true}, initialize  \tcode{exit_function} with \tcode{move(f)} otherwise initialize \tcode{exit_function} with \tcode{f}. If the copying throws an exception, calls \tcode{f()}.

\pnum
\throws Any exception thrown during the copying of \tcode{f}.

\end{itemdescr}

\begin{itemdecl}
template<class EFP>
explicit
scope_success(EFP&& f) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\remarks This constructor shall not participate in overload resolution unless \tcode{is_constructible_v<EF,EFP>} is \tcode{true}.

\pnum
\requires Given an lvalue \tcode{f} of type \tcode{EFP}, the expression \tcode{f()} shall be well formed and shall have well-defined behavior.

\pnum
\effects If \tcode{EFP} is not a lvalue-reference type and \tcode{is_nothrow_constructible_v<EF,EFP>} is \tcode{true}, initialize  \tcode{exit_function} with \tcode{move(f)} otherwise initialize \tcode{exit_function} with \tcode{f}. \\
\begin{note}
If copying fails, \tcode{f()} won't be called.
\end{note}

\pnum
\throws Any exception thrown during the copying of \tcode{f}.
\end{itemdescr}

\begin{itemdecl}
@\tcode{\textit{scope_guard}}@(@\tcode{\textit{scope_guard}}@&& rhs) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects %Copies the release state from \tcode{rhs}.
\tcode{execute_on_destruction} yields the value \tcode{rhs.execute_on_destruction} yielded before the construction.
If \tcode{is_nothrow_move_constructible_v<EF>} move constructs otherwise copy constructs \tcode{exit_function} from \tcode{rhs.exit_function}. If construction succeeds, call \tcode{rhs.release()}.

\pnum
\throws Any exception thrown during the copying of \tcode{rhs.exit_function}.
\end{itemdescr}

\begin{itemdecl}
~scope_exit() noexcept(true);
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects 
\begin{codeblock}
if (execute_on_destruction)
	exit_function();
\end{codeblock}
%Calls \tcode{exit_function()}, unless \tcode{release()} was previously called.
\end{itemdescr}

\begin{itemdecl}
~scope_fail() noexcept(true);
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects 
\begin{codeblock}
if (execute_on_destruction
   && uncaught_exceptions() > uncaught_on_creation)
	exit_function();
\end{codeblock}
%Calls \tcode{exit_function()} if its scope is left with a new exception, unless \tcode{release()} was previously called.
\end{itemdescr}

\begin{itemdecl}
~scope_success() noexcept(noexcept(exit_function()));
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects
\begin{codeblock}
if (execute_on_destruction 
   && uncaught_exceptions() <= uncaught_on_creation)
	exit_function();   
\end{codeblock}
% Calls \tcode{exit_function()} if its scope is left without an exception, unless \tcode{release()} was previously called.

\pnum
\remarks
If \tcode{noexcept(exit_function())} is \tcode{false}, \tcode{exit_function()} may throw an exception, notwithstanding the restrictions of [res.on.exception.handling].
%(c)Daniel Kruegler

\pnum
\throws
If \tcode{noexcept(exit_function())} is \tcode{false}, throws any exception thrown by \tcode{exit_function()}.

\end{itemdescr}

\begin{itemdecl}
void release() noexcept;
\end{itemdecl}

\begin{itemdescr}
\pnum
%Prevents \tcode{exit_function()} from being called on destruction.\\
\tcode{execute_on_destruction=false;}
\end{itemdescr}

\newpage

%\rSec2[scope.make_scope_exit]{\tcode{\textit{scope_guard}} factory functions}
\subsection {Scope guard factory functions [scope.make_scope_exit]}
\begin{note}
These factory functions are meant for C++14 backward compatibility of this feature. 
\end{note}

\pnum
The scope guard factory functions create \tcode{scope_exit}, \tcode{scope_fail}, and \tcode{scope_success} objects that for the function object \tcode{exit_function} evaluate \tcode{exit_function()} at their destruction unless \tcode{release()} was called.

\begin{itemdecl}
template <class EF>
scope_exit<decay_t<EF>> make_scope_exit(EF&& exit_function) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\returns \tcode{scope_exit<decay_t<EF>>(forward<EF>(exit_function));}
\end{itemdescr}

\begin{itemdecl}
template <class EF>
scope_fail<decay_t<EF>> make_scope_fail(EF&& exit_function) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\pnum
\returns \tcode{scope_fail<decay_t<EF>>(forward<EF>(exit_function));}
\end{itemdescr}

\begin{itemdecl}
template <class EF>
scope_success<decay_t<EF>> make_scope_success(EF&& exit_function) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\pnum
\returns \tcode{scope_success<decay_t<EF>>(forward<EF>(exit_function));}

\end{itemdescr}


\newpage
%%%--------- unique_resource

%\rSec2[scope.unique_resource]{Unique resource wrapper}
\subsection{Unique resource wrapper [scope.unique_resource]}

%\rSec2[scope.unique_resource.class]{Class template \tcode{unique_resource}}
\subsection {Class template \tcode{unique_resource} [scope.unique_resource.class]}

\begin{codeblock}
template<class R,class D>
class unique_resource {
public:
  template<class RR, class DD>
  explicit unique_resource(RR&& r, DD&& d)
    noexcept(is_nothrow_constructible_v<R, RR>&&
             is_nothrow_constructible_v<D, DD>)
  unique_resource(unique_resource&& rhs)
    noexcept(is_nothrow_move_constructible_v<R> &&
             is_nothrow_move_constructible_v<D>);
  unique_resource(const unique_resource&)=delete; 
  ~unique_resource();
  unique_resource& operator=(unique_resource&& rhs) ;
  unique_resource& operator=(const unique_resource&)=delete;
  void swap(unique_resource& other);
  void reset();
  template<class RR>
  void reset(RR&& r);
  void release() noexcept;
  const R& get() const noexcept;
  R operator->() const noexcept;
  @\seebelow@ operator*() const noexcept;
  const D& get_deleter() const noexcept;
private:
  R resource; // exposition only
  D deleter; // exposition only
  bool execute_on_destruction{true}; // exposition only
};
\end{codeblock}

\pnum
\begin{note}
\tcode{unique_resource} is meant to be a universal RAII wrapper for resource handles provided by an operating system or platform.
Typically, such resource handles are of trivial type and come with a factory function and a clean-up or deleter function that do not throw exceptions.
The clean-up function together with the result of the factory function is used to create a \tcode{unique_resource} variable, that on destruction will call the clean-up function. Access to the underlying resource handle is achieved through \tcode{get()} and in case of a pointer type resource through a set of convenience pointer operator functions.
\end{note}


\pnum 
The template argument
\tcode{D} shall be a 
%CopyConstructible (Table~21
%\ref{copyconstructible}
%) and 
Destructible 
(Table~27
%\ref{destructible}
) function object type~(23.14
%\ref{function.objects}
), 
for which, given
a value \tcode{d} of type \tcode{D} and a value
\tcode{r} of type \tcode{R}, the expression
\tcode{d(r)} shall be well formed, shall have well-defined behavior, and shall not throw an exception.

\pnum
\tcode{R} shall be a
%a CopyConstructible (Table~21
%\ref{copyconstructible}
%), CopyAssignable(Table~23
%\ref{copyassignable}
%), and  
Destructible 
(Table~27
%\ref{destructible}
) object type 
or a lvalue reference type.

\pnum
\begin{note} In case of \tcode{R} being a lvalue reference, an implementation should chose \tcode{reference_wrapper<R>} as type for \tcode{resource} and adjust functionality accordingly by delegating to \tcode{resource.get()} in appropriate places. In case of \tcode{D} being a \tcode{reference_wrapper} instantiation no special delegation is required, since \tcode{reference_wrapper} already delegates the call operator ([refwrap.invoke]).
\end{note} %PS: might eliminate that note again.
%, which will be wrapped with a \tcode{reference_wrapper<R>} internally.%%need to disucss
%and shall not throw an exception.  %% thanks to Daniel
%Copy construction and move assignment of \tcode{D} and \tcode{R} shall not throw an exception.

\pnum
\requires
\tcode{(is_copy_constructible_v<R> || is_nothrow_move_constructible_v<R>)}\\
\tcode{\&\& (is_copy_constructible_v<D> || is_nothrow_move_constructible_v<D>)}



\begin{itemdecl}
    template<class RR, class DD>
    explicit unique_resource(RR&& r, DD&& d)
        noexcept(is_nothrow_constructible_v<R, RR> 
                && is_nothrow_constructible_v<D, DD>)
\end{itemdecl}

\begin{itemdescr}
\pnum
% taken from Eric Niebler's code
\remarks
given 
\begin{codeblock}
template<class T, class TT>
using @\tcode{\textit{is_ntmocp_constructible}}@ =
     conditional_t<
        is_reference_v<TT> || !is_nothrow_move_constructible_v<TT>,
        is_constructible<T, TT const &>,
        is_constructible<T, TT>>;
template<class T, class TT>
constexpr auto @\tcode{\textit{is_nothrow_move_or_copy_constructible_from_v}}@=
		@\tcode{\textit{is_ntmocp_constructible}}@<T,TT>::value;
\end{codeblock}
this constructor shall not participate in overload resolution unless
\\ \tcode{\textit{is_nothrow_move_or_copy_constructible_from_v}<R, RR>} is \tcode{true}\\
and\\
\tcode{\textit{is_nothrow_move_or_copy_constructible_from_v}<D, DD>} is \tcode{true}.

\pnum
\effects If RR is not a lvalue-reference and \tcode{is_nothrow_constructible<R,RR>} is \tcode{true}, initializes \tcode{resource} with \tcode{move(r)}, otherwise initializes \tcode{resource} with \tcode{r}. Then, if \tcode{DD} is not an lvalue reference and \tcode{is_nothrow_constructible<D,DD>} is \tcode{true}, initializes  \tcode{deleter} with \tcode{move(d)}, otherwise initializes \tcode{deleter} with \tcode{d}. If construction of \tcode{resource} throws an exception, calls \tcode{d(r)}.  If construction of \tcode{deleter} throws an exception, calls \tcode{d(resource)}. 
\begin{note}
The explained mechanism should ensure no leaking resources.
\end{note}

\pnum
\throws any exception thrown during construction.
\end{itemdescr}


\begin{itemdecl}
unique_resource(unique_resource&& rhs) 
  noexcept(is_nothrow_move_constructible_v<R> &&
           is_nothrow_move_constructible_v<D>);
\end{itemdecl}
\begin{itemdescr}
\pnum
\effects If \tcode{is_nothrow_move_constructible_v<R>} is \tcode{true}, initialize \tcode{resource} from \tcode{forward<R>(rhs.resource)}, otherwise initialize \tcode{resource} from \tcode{rhs.resource}.  
\begin{note}
If construction of \tcode{resource} throws an exception \tcode{rhs} is left owning the resource and will free it in due time.
\end{note}
Then if \tcode{is_nothrow_move_constructible_v<D>} is \tcode{true} initialize \tcode{deleter} from \tcode{forward<D>(rhs.deleter)}, otherwise initialize \tcode{deleter} from \tcode{rhs.deleter}. 
If construction of \tcode{deleter} throws an exception: if \tcode{!is_nothrow_move_constructible_v<R>}, then \tcode{rhs.deleter(resource) ; rhs.release();} otherwise \tcode{rhs.resource} and \tcode{rhs.deleter} are unmodified and \tcode{rhs} can be left owning the resource. 
Finally, \tcode{execute_on_destruction} is initialized with \tcode{exchange(rhs.execute_on_destruction,false)}.
\begin{note}
The explained mechanism should ensure no leaking resources.
\end{note}
\end{itemdescr}

\begin{itemdecl}
unique_resource& operator=(unique_resource&& rhs) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\requires \\\tcode{(is_nothrow_move_assignable_v<R> || is_copy_assignable_v<R>)}
and\\
\tcode{(is_nothrow_move_assignable_v<D> || is_copy_assignable_v<D>)}

\pnum
% move(rhs.resource) vs. forward<R>(rhs.resource)
\effects If \tcode{this == \&rhs} no effect. Otherwise \tcode{reset();} followed by
Given the members \tcode{resource} and \tcode{deleter} 
If \tcode{nothrow_move_assignable_v<R>}, try to copy or move assign \tcode{deleter} from \tcode{rhs.deleter} first then \tcode{resource=forward<R>(rhs.resource)}, if  \tcode{nothrow_move_assignable_v<D>}, try to copy or move \tcode{resource} from \tcode{rhs.resource} first then \tcode{deleter=forward<D>(rhs.deleter)}, otherwise try to copy the two members. Then \tcode{execute_on_destruction = exchange(rhs.execute_on_destruction, false);}. 
\begin{note}
If a copy of a member throws an exception this mechanism leaves \tcode{rhs} intact and \tcode{*this} in the released state.
\end{note}

\pnum
\throws Any exception thrown during a copy-assignment of a member that can not be moved without an exception.
\end{itemdescr}

\pnum
\begin{note} The move semantics differ from \tcode{pair} [pairs.pair] and \tcode{tuple}[tuple.tuple], because it is important that \tcode{unique_resource} will not leak its held resource.
\end{note}.

\begin{itemdecl}
~unique_resource();
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects \tcode{reset()}.
\end{itemdescr}

\begin{itemdecl}
void reset();
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects Equivalent to
\begin{codeblock}
  if (execute_on_destruction) {
    execute_on_destruction=false;
    get_deleter()(resource);
  }
\end{codeblock}
\end{itemdescr}

\begin{itemdecl}
template <class RR>
void reset(RR && r) ;
\end{itemdecl}

\begin{itemdescr}
\pnum
Given
\begin{codeblock}
template<class T>
constexpr conditional_t<
    (!is_nothrow_move_assignable_v<T> &&
      is_copy_assignable_v<T>),
    T const &,
    T &&>
@\tcode{\textit{move_assign_if_noexcept}}@(T &x) noexcept
{
    return std::move(x);
}
\end{codeblock}

\pnum
\remarks This function shall not participate in overload resolution if \tcode{resource = \textit{move_assign_if_noexcept}(r)} is ill-formed.

\pnum
\effects 
Equivalent to
\begin{codeblock}
  reset();
  resource = @\tcode{\textit{move_assign_if_noexcept}}@(r);
  execute_on_destruction = true;
\end{codeblock}
If copy-assignment of \tcode{resource} throws an exception, \tcode{get_deleter()(r)}.
% and \tcode{release()}. %% that seems implicit from reset()
\end{itemdescr}



%Invokes the deleter function for resource if it was not previously released, e.g. \tcode{reset(); }  Then moves newresource into the tracked resource member, e.g. \tcode{resource = std::move(newresource);}  Finally sets the object in the non-released state so that the deleter function will be invoked on destruction if \tcode{release()} is not called first.

\begin{itemdecl}
void release() noexcept;
\end{itemdecl}

\begin{itemdescr}
\pnum
\effects \tcode{execute_on_destruction = false}.
\end{itemdescr}


\begin{itemdecl}
const R& get() const noexcept ;
R operator->() const noexcept ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\remarks Member
\tcode{operator->} shall not participate in overload resolution unless \\
\tcode{is_pointer_v<R> \&\& is_nothrow_copy_constructible_v<R>}\\
\tcode{\&\&(is_class_v<remove_pointer_t<R>> || }\tcode{is_union_v<remove_pointer_t<R>>)} is \tcode{true}. 

\pnum
\returns \tcode{resource}.
\end{itemdescr}

\begin{itemdecl}
@\seebelow@ operator*() const noexcept ;
\end{itemdecl}

\begin{itemdescr}
\pnum
\remarks Member \tcode{operator*} shall not participate in overload resolution unless \tcode{is_pointer_v<R>} is \tcode{true}.

\pnum
\returns \tcode{*resource}. The return type is equivalent to 
\tcode{add_lvalue_reference_t<remove_pointer_t<R>>}. 
\end{itemdescr}


\begin{itemdecl}
const D & get_deleter() const noexcept;
\end{itemdecl}

\begin{itemdescr}
\pnum
\returns \tcode{deleter}
\end{itemdescr}

%\rSec2[scope.make_unique_resource]{Factories \tcode{scope.make_unique_resource}}
\subsection {Factories for \tcode{unique_resource} [scope.make_unique_resource]}

\begin{note}
The make_unique_resource factory functions can be omitted for putting it into the standard. However, they could be useful for the LFTS3 to allow backporting to C++14.
\end{note}



\begin{itemdecl}
template<class R,class D, class S=R>
unique_resource<decay_t<R>,decay_t<D>>
make_unique_resource_checked(R&& r, S const & invalid, D && d ) 
noexcept(is_nothrow_constructible_v<decay_t<R>, R> &&
         is_nothrow_constructible_v<decay_t<D>, D>);
\end{itemdecl}

\begin{itemdescr}
\pnum
\requires 
If \tcode{s} denotes a (possibly const) value of type \tcode{S} and \tcode{r} denotes a
(possibly const) value of type \tcode{R}, the expressions \tcode{s == r} and \tcode{r == s}
are both valid, both have the same domain, both have a type that is
convertible to \tcode{bool}, and \tcode{bool(s == r) == bool(r == s)} for every \tcode{r} and
\tcode{s}. If \tcode{S} is the same type as \tcode{R}, \tcode{R} shall be EqualityComparable(Table~17
%\ref{equalitycomparable}
). 

\pnum
\effects As if
\begin{codeblock}
  bool mustrelease = bool(r == invalid);
  auto ur= unique_resource(forward<R>(r), forward<D>(d));
  if(mustrelease) ur.release();
  return ur;
\end{codeblock}

\pnum
\begin{note}
This factory function exists to avoid calling a deleter function with an invalid argument. The following example shows its use to avoid calling \tcode{fclose} when \tcode{fopen} failed and returned \tcode{NULL}.
\end{note}

\pnum
\begin{example}
\begin{codeblock}
		auto file = make_unique_resource_checked(
		      ::fopen("potentially_nonexisting_file.txt", "r"), 
		      (FILE*) NULL, &::fclose);
\end{codeblock}
\end{example}


\end{itemdescr}

\begin{note}
The following functions are optional and could be omitted for C++next.
\end{note}

\begin{itemdecl}
template<class R,class D>
unique_resource<decay_t<R>, decay_t<D>>
make_unique_resource( R && r, D && d) 
noexcept(is_nothrow_constructible_v<decay_t<R>, R> &&
         is_nothrow_constructible_v<decay_t<D>, D>);
\end{itemdecl}

\begin{itemdescr}
\pnum
\returns \\\tcode{unique_resource<decay_t<R>, decay_t<D>>(forward<R>(r), forward<D>(d))}
\end{itemdescr}


\begin{itemdecl}
template<class R,class D>
unique_resource<R&,decay_t<D>>
make_unique_resource( reference_wrapper<R> r, D d) 
noexcept(is_nothrow_constructible_v<decay_t<D>, D>);
\end{itemdecl}

\begin{itemdescr}
\pnum
\returns \tcode{unique_resource<R\&,decay_t<D>>(r.get(),forward<D>(d))}

\pnum \begin{note} There is no need to overload on \tcode{reference_wrapper<D>} for the deleter. \end{note}
\end{itemdescr}


\newpage
\chapter{Appendix: Example Implementations}
removed, see \\
https://github.com/PeterSommerlad/SC22WG21_Papers/tree/master/workspace/P0052_scope_exit/src
\end{document}

