\documentclass[ebook,11pt,article]{memoir}
\usepackage{geometry}  % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}  % ... or a4paper or a5paper or ... 
%\geometry{landscape}  % Activate for for rotated page geometry
%%% from std.tex
%\usepackage[american]
%           {babel}        % needed for iso dates
%\usepackage[iso,american]
%           {isodate}      % use iso format for dates
\usepackage[final]
           {listings}     % code listings
\usepackage{longtable}    % auto-breaking tables
\usepackage{ltcaption}    % fix captions for long tables
\usepackage{relsize}      % provide relative font size changes
%\usepackage{textcomp}     % provide \text{l,r}angle
\usepackage{underscore}   % remove special status of '_' in ordinary text
%\usepackage{parskip}      % handle non-indented paragraphs "properly"
%\usepackage{array}        % new column definitions for tables
\usepackage[normalem]{ulem}
\usepackage{enumitem}
\usepackage{color}        % define colors for strikeouts and underlines
%\usepackage{amsmath}      % additional math symbols
%\usepackage{mathrsfs}     % mathscr font
\usepackage[final]{microtype}
%\usepackage{multicol}
\usepackage{xspace}
%\usepackage{lmodern}
\usepackage[T1]{fontenc} % makes tilde work! and is better for umlauts etc.
%\usepackage[pdftex, final]{graphicx}
\usepackage[pdftex,
%            pdftitle={C++ International Standard},
%            pdfsubject={C++ International Standard},
%            pdfcreator={Richard Smith},
            bookmarks=true,
            bookmarksnumbered=true,
            pdfpagelabels=true,
            pdfpagemode=UseOutlines,
            pdfstartview=FitH,
            linktocpage=true,
            colorlinks=true,
            linkcolor=blue,
            plainpages=false
           ]{hyperref}
%\usepackage{memhfixc}     % fix interactions between hyperref and memoir
%\usepackage[active,header=false,handles=false,copydocumentclass=false,generate=std-gram.ext,extract-cmdline={gramSec},extract-env={bnftab,simplebnf,bnf,bnfkeywordtab}]{extract} % Grammar extraction
%
\renewcommand\RSsmallest{5.5pt}  % smallest font size for relsize


%%%% reuse all four from std.tex:
\input{layout}
\input{styles}
\input{macros}
\input{tables}

\renewcommand{\iref}[1]{[#1]}


\pagestyle{myheadings}

\newcommand{\papernumber}{d1411r1}
\newcommand{\paperdate}{2019-02-20}

\markboth{\papernumber{} \paperdate{}}{\papernumber{} \paperdate{}}

\title{\papernumber{} - Please reconsider <scope> for C++20}
\author{Peter Sommerlad}
\date{\paperdate}                % Activate to display a given date or no date
\setsecnumdepth{subsection}

\begin{document}
\maketitle
\begin{center}
\begin{tabular}[t]{|l|l|}\hline 
Document Number:&  \papernumber \hfill \\\hline
Date: & \paperdate \\\hline
Project: & Programming Language C++\\\hline 
Audience: & LEWG\\\hline
Target: & C++20 \\\hline
\end{tabular}
\end{center}
\chapter{Introduction}
Timing is hard. When LEWG blessed p0052 (scope guards and unique_resource) to advance (see \url{https://issues.isocpp.org/show_bug.cgi?id=6}), neither C++20 nor a library fundamentals TS was open and the paper was not ready for C++17. The current working draft was for C++17 and the library fundamentals TS 2 was (about to be) published. Thinking ahead LEWG voted p0052 for a future library fundamentals TS 3. I believe that went under LEWG's assumption that such a vehicle would open early enough so that the highly desired and long brewing feature could be included into C++20. However, library fundamentals TS 3 was only opened for business summer 2018 and thus too late to include new features as a staging area for C++20. While getting improved through LWG feedback under the assumption it could and should be included into C++20, it turned out, that such a formal blessing by LEWG is missing. In addition some minor design question arose (default constructability of unique_resource) that must be addressed by LEWG. Many people have expressed the desire to get p0052 into C++20, or at least its unique_resource part.

To not overwhelm LEWG with the wording and rationale of p0052 I would like to ask for answering 2-3 simple questions as soon as possible, so that I can be relieved of p0052 (which is reaching an age it has to go to elementary school). I understand that LEWG does not bless for inclusion in the working draft, but we need a formal forward decision by LEWG to not get bailed with that paper when it comes to plenary.

\added{FYI: LWG blessed p0052 wording (including the default ctor) in Kona, it is just up to LEWG if it will go into the IS for C++20.}


\chapter{Decisions to be made}

\section{Decision 1: Allow default constructability of unique_resource}
I have user requests that desire \tcode{unique_resource} to provide a default constructor if its resource and deleter types are default constructible. The reason is to ease using member variables and containers of such unique_resource objects. The default constructor would create the unique_resource in a released state to be later reassigned or reset. This would not introduced additional overhead, since the underlying infrastructure must already deal with such a released state.

LEWG question: Should unique_resource provide a default constructor creating a resource in a released state if both template argument types allow so?

\added{
For convenience, here is the specification to be added to [scope.unique_resource.ctor].
}

\begin{addedblock}
\begin{itemdecl}
unique_resource()
\end{itemdecl}

\begin{itemdescr}
\pnum
\constraints
\tcode{is_default_constructible_v<R>} \tcode{\&\&}
\tcode{is_default_constructible_v<D>} is \tcode{true}.

\pnum
\effects
Value-initializes
\tcode{resource} and \tcode{deleter}; \tcode{execute_on_reset} is initialized with \tcode{false}.
\end{itemdescr}
\end{addedblock}



\section{Decision 2: Bless all of p0052 forward to LWG to consider for inclusion in C++20}

LEWG question: Should p0052 be forwarded for C++20? (it was in spirit, but not in fact).

\section{Decision 3 (only if 2 is NO): Bless unique_resource part of p0052 to LWG to consider for inclusion in C++20}

This question is only relevant if Decision 2 is NO. There is stronger user desire for unique_resource than for scope guards.

LEWG question: Should only the unique_resource part of p0052 be forwarded to LWG for C++20?


\section{Outlook}
If neither decision 2 or 3 are positive I am eagerly looking for a future champion and co-author that takes this paper further. I have other priorities to work on and am burnt out by the many iterations this paper took.

TODO

history: started Andrew Sandoval via email to std-proposals Nov 4 2012.
missed C++14,C++17 and C++20...

volatility of LEWG broke it.

\end{document}

