\documentclass[ebook,11pt,article]{memoir}
\usepackage{geometry}  % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}  % ... or a4paper or a5paper or ... 
%\geometry{landscape}  % Activate for for rotated page geometry
%%% from std.tex
%\usepackage[american]
%           {babel}        % needed for iso dates
%\usepackage[iso,american]
%           {isodate}      % use iso format for dates
\usepackage[final]
           {listings}     % code listings
\usepackage{longtable}    % auto-breaking tables
\usepackage{ltcaption}    % fix captions for long tables
\usepackage{relsize}      % provide relative font size changes
%\usepackage{textcomp}     % provide \text{l,r}angle
\usepackage{underscore}   % remove special status of '_' in ordinary text
%\usepackage{parskip}      % handle non-indented paragraphs "properly"
%\usepackage{array}        % new column definitions for tables
\usepackage[normalem]{ulem}
\usepackage{enumitem}
\usepackage{color}        % define colors for strikeouts and underlines
%\usepackage{amsmath}      % additional math symbols
%\usepackage{mathrsfs}     % mathscr font
\usepackage[final]{microtype}
%\usepackage{multicol}
\usepackage{xspace}
%\usepackage{lmodern}
\usepackage[T1]{fontenc} % makes tilde work! and is better for umlauts etc.
%\usepackage[pdftex, final]{graphicx}
\usepackage[pdftex,
%            pdftitle={C++ International Standard},
%            pdfsubject={C++ International Standard},
%            pdfcreator={Richard Smith},
            bookmarks=true,
            bookmarksnumbered=true,
            pdfpagelabels=true,
            pdfpagemode=UseOutlines,
            pdfstartview=FitH,
            linktocpage=true,
            colorlinks=true,
            linkcolor=blue,
            plainpages=false
           ]{hyperref}
%\usepackage{memhfixc}     % fix interactions between hyperref and memoir
%\usepackage[active,header=false,handles=false,copydocumentclass=false,generate=std-gram.ext,extract-cmdline={gramSec},extract-env={bnftab,simplebnf,bnf,bnfkeywordtab}]{extract} % Grammar extraction
%
\renewcommand\RSsmallest{5.5pt}  % smallest font size for relsize


%%%% reuse all four from std.tex:
\input{layout}
\input{styles}
\input{macros}
\input{tables}

\renewcommand{\iref}[1]{[#1]}


\pagestyle{myheadings}

\newcommand{\papernumber}{p1410r0}
\newcommand{\paperdate}{2019-01-21}

\markboth{\papernumber{} \paperdate{}}{\papernumber{} \paperdate{}}

\title{\papernumber{} - Remove deprecated strstream}
\author{Peter Sommerlad}
\date{\paperdate}                % Activate to display a given date or no date
\setsecnumdepth{subsection}

\begin{document}
\maketitle
\begin{center}
\begin{tabular}[t]{|l|l|}\hline 
Document Number:&  \papernumber \hfill \\\hline
Date: & \paperdate \\\hline
Project: & Programming Language C++\\\hline 
Audience: & LEWG/LWG\\\hline
Target: & C++20 or C++23\\\hline
\end{tabular}
\end{center}
\chapter{Introduction}
Streams have been the oldest part of the C++ standard library and especially strstreams that can use pre-allocated buffers have been deprecated for a long time now, waiting for a replacement. p0407, p0408 and p0448 provide together a modern complete replacement for strstream. This paper provides wording for actually removing it from appendix D and referring to its removal in appendix C. If p0408 (that includes p0407) and p0448 are accepted for C++20 then this paper can either be applied for C++20 or as some propose applied for C++23. I personally believe time is ripe for C++20, but I let LEWG or the full committee decide.

\chapter{Impact on the standard}
This section is relative to n4791 which is current at the time of the writing.

\section{provide wording for difference in Appendix C}
Depending on approval standard add the following paragraph to section [diff.cpp17.depr] (or [diff.cpp20.depr] when that exists). The editorial guidance now is relative to n4791.

Add a new paragraph in [diff.cpp17.depr] before p.7 as follows:

\nodiffref
\change Remove deprecated \tcode{char *} streams 
\rationale
The introduction of span streams [span.streams] and the provisioning of efficient access to the underlying buffer of string streams [string.streams] provide better alternatives for the use cases of \tcode{char *} streams. The latter have been deprecated in all revisions of this international standard.
\effect
A valid \CppXVII{} program that relies on the deprecated \tcode{char *} stream classes may fail to compile.
In cases where the underlying \tcode{strstreambuf} is allows writing and its \tcode{strstate} bitmask has \tcode{allocated} or \tcode{dynamic} set, classes from [string.streams] should be used instead. In other cases either classes from [span.streams] or [string.streams] can be used to achieve its functionality with equivalent performance.

\section{Remove deprecated \tcode{char *} streams from Appendix D}
Remove all of Appendix D.8  [depr.str.strstreams]


\end{document}

